Spatial clustering and violent crime near places of worship in Recife, Brazil. Do moral communities have a spatial dimension?

### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime.Rmd
Revisions before submission to Journal of Quantitative Criminology

---
title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
  Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

@@

Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral]. It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential. The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms in a consistent way - as is the case in Brazil [@Garmany2014space].

@@

In what follows, we propose to adopt a disaggregated approach by considering the intensity of criminal events with respect to places of worship. This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.

@@

As noted above, previous research that has investigated the presence of churches from a geographical perspective has used data aggregated at different scales. In the case of Nie and Yang [-@Nie2019smoking], the unit of analysis was the county, whereas Warner and Konkel [-@Warner2019neighborhood] used Census Block Groups, a much smaller unit of analysis. Research in spatial criminology, on the other hand, has studied criminal events as point patterns since at least the work of Levine et al. [-@Levine1986crime] investigated the concentration of criminal events in the proximity of bus stops. Since then, many other works have applied tools of spatial point pattern analysis to investigate the empirical properties of the distribution of crime. This includes Craglia et al. [-@Craglia2000comparative] who used the Ripley's K-function [@Ripley1976second] to investigate clustering processes, and Rogerson and Sun [-@Rogerson2001spatial] who applied nearest neighbor techniques to the study of arson. More recent studies of crime as point patterns are Nakaya and Yano [-@Nakaya2010visualizing], Kiani et al. [-@Kiani2015analysis], and Malleson and Andresen [-@Malleson2015spatio].

@@

The data are organized at the individual level and it is possible to obtain information about the crime event location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim. The database used in this study comprises the period from July 1, 2008 to June 30, 2010. Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of LIVC in the period analyzed were men. In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime. Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms. Figure \ref{fig:plot-crime} shows the spatial distribution of the 1,657 LIVC crimes that occurred in the city of Recife between July, 2008 and June, 2010.

>>>>

---
title: Spatial clustering and violent crime near places of worship in Recife, Brazil. Do moral communities have a spatial dimension?
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
  Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

@@

Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral]. It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential. The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms in a consistent way, such as disadvanteged Black communities in the American South [e.g., @Thomson2020religion] and in regions in South America such as Brazil [e.g., @Garmany2014space].

@@

In what follows, we propose to adopt a disaggregated approach by considering the intensity of criminal events with respect to places of worship. An advantage of a disaggregated approach is the use of micro-spatial units to investigate the spatial concentration (or clustering) of events [see @Andresen2017crime]. This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.

@@

As noted above, previous research that has investigated the presence of churches from a geographical perspective has used data aggregated at different scales. In the case of Nie and Yang [-@Nie2019smoking], the unit of analysis was the county, whereas Warner and Konkel [-@Warner2019neighborhood] used Census Block Groups, a much smaller unit of analysis. Research in spatial criminology, on the other hand, has studied criminal events as point patterns since at least the work of Levine et al. [-@Levine1986crime] investigated the concentration of criminal events in the proximity of bus stops. Since then, many other works have applied tools of spatial point pattern analysis to investigate the empirical properties of the distribution of crime. This includes Craglia et al. [-@Craglia2000comparative] who used the Ripley's K-function [@Ripley1976second] to investigate clustering processes, and Rogerson and Sun [-@Rogerson2001spatial] who applied nearest neighbor techniques to the study of arson. More recent studies of crime as point patterns are Nakaya and Yano [-@Nakaya2010visualizing], Kiani et al. [-@Kiani2015analysis], and Malleson and Andresen [-@Malleson2015spatio], .

@@

The data are organized at the individual level and it is possible to obtain information about the crime event location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim. The database used in this study comprises the period from July 1, 2008 to June 30, 2010. Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of LIVC in the period analyzed were men. In addition, most of the victims were Black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime. Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms. Figure \ref{fig:plot-crime} shows the spatial distribution of the 1,657 LIVC crimes that occurred in the city of Recife between July, 2008 and June, 2010.

### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime.tex
Revisions before submission to Journal of Quantitative Criminology

\title{Do moral communities have a spatial dimension? A spatial exploratory
analysis of places of worship and violent crime in the city of Recife,
Brazil \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

@@

Yet another fruitful avenue for research, and one that has only recently
begun to be explored, is the presence of environmental attributes that
can help to reinforce moral norms, such as schools and churches (e.g.,
Abdullah et al. 2018; Davignon and Thomson 2015; Furr-Holden et al.
2010; Traunmuller 2011). It is thus that in a recent paper, Warner and
Konkel (2019) note that the role of places of worship, as distinct
entities from the members of the congregations, have received less
attention in empirical and theoretical research for their deterrence
potential. The role of these institutions might be particularly
important in places where formal state institutions lack the means or
the will to enforce norms in a consistent way - as is the case in Brazil
(Garmany 2014).

@@

In what follows, we propose to adopt a disaggregated approach by
considering the intensity of criminal events with respect to places of
worship. This requires micro-level data both of crimes and places of
worship, and the use of appropriate spatial analytical tools for point
patterns, as discussed next.

@@

As noted above, previous research that has investigated the presence of
churches from a geographical perspective has used data aggregated at
different scales. In the case of Nie and Yang (2019), the unit of
analysis was the county, whereas Warner and Konkel (2019) used Census
Block Groups, a much smaller unit of analysis. Research in spatial
criminology, on the other hand, has studied criminal events as point
patterns since at least the work of Levine et al.~(1986) investigated
the concentration of criminal events in the proximity of bus stops.
Since then, many other works have applied tools of spatial point pattern
analysis to investigate the empirical properties of the distribution of
crime. This includes Craglia et al.~(2000) who used the Ripley's
K-function (Ripley 1976) to investigate clustering processes, and
Rogerson and Sun (2001) who applied nearest neighbor techniques to the
study of arson. More recent studies of crime as point patterns are
Nakaya and Yano (2010), Kiani et al.~(2015), and Malleson and Andresen
(2015).

@@

The data are organized at the individual level and it is possible to
obtain information about the crime event location, day of the week, day
of the month, period of the day, as well as gender, age, and race of the
victim. The database used in this study comprises the period from July
1, 2008 to June 30, 2010. Some descriptive statistics regarding this
dataset are reported in Table 1, where it can be seen that overall,
about 91\% of the victims of LIVC in the period analyzed were men. In
addition, most of the victims were black or brown, and the youth
population between the ages of 18 and 30 is the most affected by violent
crime. Lastly, it should be noted that about 88\% of the criminal events
under analysis involved firearms. Figure \ref{fig:plot-crime} shows the
spatial distribution of the 1,657 LIVC crimes that occurred in the city
of Recife between July, 2008 and June, 2010.

@@

@@

>>>>

\title{Spatial clustering and violent crime near places of worship in Recife,
Brazil. Do moral communities have a spatial dimension? \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

@@

Yet another fruitful avenue for research, and one that has only recently
begun to be explored, is the presence of environmental attributes that
can help to reinforce moral norms, such as schools and churches (e.g.,
Abdullah et al. 2018; Davignon and Thomson 2015; Furr-Holden et al.
2010; Traunmuller 2011). It is thus that in a recent paper, Warner and
Konkel (2019) note that the role of places of worship, as distinct
entities from the members of the congregations, have received less
attention in empirical and theoretical research for their deterrence
potential. The role of these institutions might be particularly
important in places where formal state institutions lack the means or
the will to enforce norms in a consistent way, such as disadvanteged
Black communities in the American South (e.g., Thomson, n.d.) and in
regions in South America such as Brazil (e.g., Garmany 2014).

@@

In what follows, we propose to adopt a disaggregated approach by
considering the intensity of criminal events with respect to places of
worship. An advantage of a disaggregated approach is the use of
micro-spatial units to investigate the spatial concentration (or
clustering) of events (see Andresen, Linning, and Malleson 2017). This
requires micro-level data both of crimes and places of worship, and the
use of appropriate spatial analytical tools for point patterns, as
discussed next.

@@

As noted above, previous research that has investigated the presence of
churches from a geographical perspective has used data aggregated at
different scales. In the case of Nie and Yang (2019), the unit of
analysis was the county, whereas Warner and Konkel (2019) used Census
Block Groups, a much smaller unit of analysis. Research in spatial
criminology, on the other hand, has studied criminal events as point
patterns since at least the work of Levine et al.~(1986) investigated
the concentration of criminal events in the proximity of bus stops.
Since then, many other works have applied tools of spatial point pattern
analysis to investigate the empirical properties of the distribution of
crime. This includes Craglia et al.~(2000) who used the Ripley's
K-function (Ripley 1976) to investigate clustering processes, and
Rogerson and Sun (2001) who applied nearest neighbor techniques to the
study of arson. More recent studies of crime as point patterns are
Nakaya and Yano (2010), Kiani et al.~(2015), and Malleson and Andresen
(2015), .

@@

The data are organized at the individual level and it is possible to
obtain information about the crime event location, day of the week, day
of the month, period of the day, as well as gender, age, and race of the
victim. The database used in this study comprises the period from July
1, 2008 to June 30, 2010. Some descriptive statistics regarding this
dataset are reported in Table 1, where it can be seen that overall,
about 91\% of the victims of LIVC in the period analyzed were men. In
addition, most of the victims were Black or brown, and the youth
population between the ages of 18 and 30 is the most affected by violent
crime. Lastly, it should be noted that about 88\% of the criminal events
under analysis involved firearms. Figure \ref{fig:plot-crime} shows the
spatial distribution of the 1,657 LIVC crimes that occurred in the city
of Recife between July, 2008 and June, 2010.

@@

\leavevmode\hypertarget{ref-Andresen2017crime}{}%
Andresen, M. A., S. J. Linning, and N. Malleson. 2017. ``Crime at Places
and Spatial Concentrations: Exploring the Spatial Stability of Property
Crime in Vancouver Bc, 2003-2013.'' Journal Article. \emph{Journal of
Quantitative Criminology} 33 (2): 255--75.
\url{https://doi.org/10.1007/s10940-016-9295-8}.

@@

\leavevmode\hypertarget{ref-Thomson2020religion}{}%
Thomson, R. A. n.d. ``Religion in Civil Society: The Influence of Black
Religious Ecology on Crime in the South.'' Journal Article.
\emph{Journal of Quantitative Criminology}, 27.
\url{https://doi.org/10.1007/s10940-019-09444-7}.

### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime_Title_Page.Rmd
Revisions before submission to Journal of Quantitative Criminology

---
title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
    **Funding:** Edivaldo Constantino das Neves Júnior received support from the Government of Canada, through the program Emerging Leaders in the Americas. **Conflicts of Interest:** The authors declare that they have no conflict of interest.

>>>>

---
title: Spatial clustering and violent crime near places of worship in Recife, Brazil. Do moral communities have a spatial dimension?
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
    **Funding:** Edivaldo Constantino das Neves Júnior received support from the Government of Canada, through the program Emerging Leaders in the Americas. **Conflicts of Interest:** The authors declare that they have no conflict of interest.

### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime_Title_Page.tex
Revisions before submission to Journal of Quantitative Criminology

\title{Do moral communities have a spatial dimension? A spatial exploratory
analysis of places of worship and violent crime in the city of Recife,
Brazil \thanks{\textbf{Funding:} Edivaldo Constantino das Neves Júnior received support
from the Government of Canada, through the program Emerging Leaders in
the Americas. \textbf{Conflicts of Interest:} The authors declare that
they have no conflict of interest.} }

>>>>

\title{Spatial clustering and violent crime near places of worship in Recife,
Brazil. Do moral communities have a spatial dimension? \thanks{\textbf{Funding:} Edivaldo Constantino das Neves Júnior received support
from the Government of Canada, through the program Emerging Leaders in
the Americas. \textbf{Conflicts of Interest:} The authors declare that
they have no conflict of interest.} }
